## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}
a <- 10
if (a <=5) {
  print(a)
} else {
  print("Its more than 5")
}
```

We can also add else if statements in the middle if we need more options.

```{r}
a <- 15
if(a <=5){
  print("small")
} else if (a > 5 & a <= 10){
  print("medium")
} else if(a > 10 | a == 7){
  print("Large")  
}else {
    print("Huge")
  }
```

Neat! We can also nest conditionals just like we do in Python.

```{r}
a <- 5
if (a / 2 != 0){
  if (a == 5){
    print("a is 5")
  }
}
    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for(value in 1:5){
  #paste is the concatenation funcion
  print(paste(value, "is greater than", value - 1))
}
```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
a_vector <- c(1,1,1,1,1,1,1)

#change the first, third, and sixth elements to the value 6

for (i in c(1,3,6)){
  a_vector[i] <- 6
}

a_vector
```

We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow = 2, ncol = 2)
for (row in 1:2){
  for (col in 1:2){
    mat[row,col] <- 0 #try changin it to mat[col,row]
    print(mat)
  }
}
```

Like Python, we have while loops in R:

```{r}
b <- 0
while (b < 10) {
  # increase b by 2
  b <- b + 2
  #print b
  print(b)
}
```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
  
}

```

An example:

```{r}
test_temp <- c(16.550540,26.942005,32.591088,8.580116,28003873,28.542391,20.976820,21.173577,21.048836,18.769735)

#then we want a function that converts from C to F:
c_to_f <- function(dat){
  temp_fahr <- dat*(9/5) + 32
  
  return(temp_fahr)
}

print(c_to_f(26))


```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- c_to_f(test_temp)
data_fahr #my answers arent matching the tutorial
```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <- matrix(c(1,1,1,2,2,2), nrow <- 2)
test_mat 
apply(test_mat, 2, sum) #1 is for rows, 2 is for columns
```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
test_temp_mat <- matrix(c(16.550540,26.942005,32.591088,8.580116,28003873,28.542391,20.976820,21.173577,21.048836,18.769735), nrow = 2)

apply(test_temp_mat,1, c_to_f)
```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
# 1). Create a function to determine what quater a month belongs to
    quarter_of <- function(month){
  quat <- c(1,2,3,4)
  
  if(month <= 3){
    print(1)}
  else if (month >= 4 & month < 7){
    print(2)}
  else if (month >=7 & month <10 ) {
    print(3)    
  }
  else {
    print (4)
  }
  
#create a function that will tell in how many years a father will be twice as old as the son
  
twice_as_old <- function(dad_years_old, son_years_old){
     twice_age <- abs(2 * son_years_old - dad_years_old)
    return(twice_age)
}

#take in 3 numbers, return the max number combindation expression_matter <- function(a, b, c) {
  # your code here
   combinations <- c(
    a + b + c,
    a + b * c,
    a * b + c,
    a * b * c,
    (a + b) * c,
    a * (b + c)
  )
  return(max(combinations))
    }
    
#Create a function to calculate grades avg based on number of projects submitted and grade on final
  final_grade <- function(exam, projects){
    if (exam > 90 || projects > 10) {
    return(100)
  } else if (exam > 75 && projects >= 5) {
    return(90)
  } else if (exam > 50 && projects >= 2) {
    return(75)
  } else {
    return(0)
  }
}  
#give a yr return the centruy its in
  century <- function(year) {
  # your code goes here
   return(ceiling(year / 100))
  }
  
#
    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
fizz_buzz <- function(number) {
  if (number %% 3 == 0 && number %% 5 == 0) {
    return("fizzbuzz")
  } else if (number %% 3 == 0) {
    return("fizz")
  } else  {
    return("buzz")
  } 
}
    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
    marioStairs <- function(height) {
  for (i in 1:height) {
    cat(rep("X", i), "\n", sep="")
  }
}


marioStairs(4)
    ```
