## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
install.packages("tidyverse")
library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
myplot <- mpg

myplot
view(myplot)
``` 



**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
mtcars

```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
#also try color = class
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = rgb(0.7,1,0.3))) +
  facet_wrap(~class, nrow = 3) #try shape = class, color = class
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes( x = displ, y = hwy, color = drv )) +
  facet_grid(drv~ .)
```

```{r}
#ggplot(data = mpg)+
  #geom_smooth(mapping = aes(x = displ, y = hwy, linetype= drv))+
  #geom_point(mapping = aes(x = displ, y = hwy, color = drv))
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv))+
  geom_point(mapping = aes(color = class))+
  geom_smooth(se = FALSE)
```

```{r}

```

```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut))
```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console.It shows up because this is a discrete dataset. Rather than the continious dataset that we have been using.

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds)+
  stat_count(mapping = aes(x = cut))
```

Let's add color here!

```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = cut))
```

What if we add to a different variable than cut?

```{r}
pie = ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = clarity), #try dodge = clarity
          position = "fill"
  )+
  theme(aspect.ratio = 1)+
  labs(x = NULL, Y = NULL)
pie + coord_polar()
```

**For your consideration...** Is the above a good graph?

```{r}
alt- becomes <- (option- on mac)
```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; Analyze that data set, create graphs, find a hypothesis for 3 data sets
View(esoph)
```


```{r}
ggplot(data = esoph)+
  geom_point(mapping = aes(x = ncases, y = agegp, color = agegp))
#my first observation is that the age gaps 25-34, and 75+ have the least amount of
#number of cases. 
```


```{r}
#view(Orange)
ggplot(data = Orange) +
  geom_point(mapping = aes( x = age, y = circumference, color = Tree )) +
  facet_grid(Tree~ .)
#as the age of the tree increases, so does the circumference of the orange
```
```{r}
#view(sleep)
ggplot(data = sleep)+
  geom_point(mapping =  aes(x= group, y = extra, color = group))

#from this plot I can tell that group 2 drug allowed patients to get the most amount of extra sleep.
```
```{r}

```
